---
title: "File_Handing"
author: "Hafez Ahmad"
date: "`r Sys.Date()`"
output: html_document
---
# Files 

**Most used Files in R**

1.  ***CSV***/***TXT***/***XLSX*** : Tabular data.
2.  ***JSON***: web applications and APIs.
3.  ***Parquet***: Apache Spark.
4.  ***SQ***L :Databases.
5.  ***Shapefile***: points, lines, and polyggonal shapes.
6.  ***GeoTIFF***: satellite imagery and digital elevation models.
7.  ***NetCDF***/***HDF5***/***ZAR***: meteorology, oceanography, and climate 
8.  ***FASTA***/***BAM***/***NEXUS***: DNA and protein sequences, as well as phylogenetic trees.

**Functions for reading data**:
```{r echo=FALSE,results='asis'}
library(knitr)
```

1.  `read.csv()`: Reads a CSV file and returns a data frame.
```{r}
df<- read.csv("E:/R/data/iris.csv")

```

## view first 3 rows head()
`head(df,3)`
```{r echo=FALSE}
kable(head(df,3))
```
## view last 2 rows tail()
`tail(df,2)`
```{r echo=FALSE}
kable(tail(df,2))
```

2.  `read.table()`: Reads a tab-delimited or fixed-width file and returns a data frame.

3.  `read.delim()`: Reads a delimited file and returns a data frame.

4.  `read.fwf():` Reads a fixed-width file and returns a data frame.

5.  `read.xlsx()`: Reads an Excel file and returns a data frame.

***Functions for writing data***:

1.  `write.csv()`: Writes a data frame to a CSV file.
`write.csv(currentdata, storefile,others)`
```{r}
write.csv(df,"E:/R/data/iris.csv",row.names = FALSE)
```


2.  `write.table()`: Writes a data frame to a tab-delimited or fixed-width file.

3.  `write.delim()`: Writes a data frame to a delimited file.

4.  `write.fwf()`: Writes a data frame to a fixed-width file.

5.  `write.xlsx()`: Writes a data frame to an Excel file.

# Some statistics
```{r}
summary(df)
```
***Calculating Column mean***
`mean(df$Sepal.Length, na.rm = TRUE)`
```{r message=FALSE, warning=FALSE}
library(dplyr)

summarize(df[1:5,], mean(Sepal.Length))
```
`df$x_mean <- mean(df$x)`
```{r}
mutate(df[1:5,], Sepal.Length_mean = mean(Sepal.Length))

```
`names(df)[names(df) == "x"] <- "x_new"`
```{r}
rename(df[1:5,], Sepal.Length_new = Sepal.Length)

```

`df[df$x >= 1, ]`
```{r}
filter(df[1:5,], Sepal.Length >= 1)

```
`aggregate(df[c("x", "y")], by = list(df$grp), mean)`

```{r}
summarize(group_by(df, Species), mean(Sepal.Length))

```

```{r}
summarize(df, across(starts_with("Sepal.Length"), mean))

```


```{r}
summarize(df, across(c(Sepal.Length, Petal.Length), list(max, min)))

```
`rbind(value = c(min(df$Sepal.Length, na.rm = TRUE), max(df$Sepal.Length, na.rm = TRUE)))`

```{r}
mutate(df[1:5,], cor(Sepal.Length, Petal.Length))

```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```

